### 2008-03-15 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Finishing section about fully specifying up front

There are generally two approaches to creating an architecture, as defined y Bass et. al. and that is to define it in advance, and only define the overall structure and then let it grow from the following implementations.

Creating an architecture only makes sense if it is believed that it may be used in other products, and the number of products in the pipeline and how well their commonalities may be predicted, helps to determine which approach to use. Naturally the situation is not black and white, as there are many possibilities between the fully evolutionary architecture to the fully initially defined, yet for the purpose of debating advantages and disadvantages the simple black and white is simply easier.

Defining the architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money. The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the architecture and very little (if any) redesign has to be done. As more and more products are created based on the architecture their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.

Letting the architecture evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product. When the second product is to be created more of the architecture is defined based on the parts of the first product, which may be used in the second. Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire architectur was defined in advance. As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.

If the architecture is used to define a Product Line it becomes even more pronounced, and the above two approaches are defined as Heavyweight and Lightweight, respectively. A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in figure xx.

\clearpage

\begin{figure}[!htb]
\centerline{\epsfig{figure=figures/product-line,scale=1.0 }}
\caption{Heavy-weight vs. Light-weight}
\label{fig:product-line}
\end{figure}

Deciding which approach to use, is more than just a matter of choosing between Heavyweight and Lightweight (or fully defined or evolutionary) based on the number of expected products, it is also a matter of how well the common architecture of the future products can be predicted, for if the architecture needs to be changed from product to product then all the advantage is lost.

Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well know that the architecture can be fully specified in advance.

>>>>

There are generally two approaches to creating an architecture, as defined by Bass et. al. and that is to define it in advance, and only define the overall structure and then let it grow from the following implementations.

There are also two reasons for creating an architecture; in order to create a product that exhibits some of the quality attributes defined in the architecture (modifiability, usability, ...) and in order to reuse the architecture (or part of the architecture) as a product line for creating other products. This latter reason is described in section yy.

Creating an architecture up front allows for the definition of all properties of the architecture; the quality attributes and their individual priorities, the individual elements/components and how they interconnect both at runtime and statically, etc. This has the benifit that when the design and implementation phase is begun it may easily be broken down into parallel activities and integrated later, as everyone has a clear picture of what is important and what the interfaces are.

Unfortunately defining the architecture fully up front only makes sense if there is enough information in advance to create the architecture form. Very often there are "dark horses", things that do not become apparent until the design and implementation phase. If it is not believed that sufficient information exist to fully define the architecture, then doing so may not only be a waste, it may lead to the very dangerous loose-loose principle, where e.g. an overhead is introduced to service modifiability, only to be hacked in order to service a performance issue that was not apparent at the architectual design pahse. This means that part of the overhead is preserved (in memory-footprint if nowhere else), without preserving any of the modifiability advantages.

Another requirement to defining the architecture fully up front has to do with the organization. It may be good and dandy to create a fully defined architecture with focus on parallel implementation, if there is only one in-house developer to implement the product.
Based on this the Beer Web Store architect must have decided that there are sufficient information to fully define the architecture and also that there is a reason (parallel implementation, out-sourceing of the design/implementation, etc.).

Asuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectual properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has.
